Supplement A to 'Researchers understand and use the logic of significance testing in a novel statistical reasoning task

### 2020-02-05 inst/manuscript/manuscript.Rmd
small changes ahead of resubmission

The interpretation of studies of researchers' understanding of SST are limited, however, by their methodology. A typical study presents a vignette describing research results. Statistical results are offered to the participants (e.g., a $t$ statistic and $p$ value), who are then asked to explicitly give or endorse various interpretations that are thought to represent participants' understanding of SST. However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication). It is unclear how well vignette studies (including ones by the present authors; @Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results. Conceptual understanding and fluency with common representations are both important, but are distinct.

@@

using a simulated experimental task, we tested researchers' and trainees' understanding and application of SST. Our key innovation allowing us to focus on SST was to prevent---or make difficult---the use of alternative strategies. If participants have poor understanding of SST, one would predict they would perform poorly and would not be sensitive to SST-related information. In fact, scientists performed well, explicitly reported using SST strategies, and were sensitive to SST-related information. Our results strongly suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.

@@

Participants were scientists or trainees recruited via social media. Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of Christmas elves --- "Jinglies" or "Sparklies" --- could make more of a particular toy, where the true effect size was a randomly-assigned group difference between 0 and 1 standard deviations ($\delta$). The numerical information for an experiment, including sample sizes and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}). Participants could change the sample size per group for each experiment (increasing the time required to return a result), but did not know its numerical value. The interpretation of these colors and locations was unknown to the participants, aside from the monotone relationship with the statistical evidence.

@@

Participants were also offered the opportunity to sample "random shuffle reports" that were described as the results of experiments with random assignment of elves to groups: that is, the result of experiments in which the null hypothesis was true. These results took no time to return. Participants were not instructed how to use these samples, only that they might use them.

@@

We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategy), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports.

@@

We determined the coding scheme and independently categorized the first 20 participant, discussing the source of disagreements. After categorizing the remaining participants, some disagreements were resolved through agreement, and a discussion between the authors was had over what caused the disagreements. The remainder of the disagreements were re-coded separately, and a final round of discussion resolved the remaining disagreements. Participants' codings described in detail in Supplement B.

@@

* **Data and materials availability**:

* Supplementary Information is linked to the online version of the paper at www.nature.com/nature

* **Author Contributions**:

* Reprints and permissions information is available at www.nature.com/reprints.

>>>>

The interpretation of studies of researchers' understanding of SST are limited, however, by their methodology. A typical study presents a vignette describing research results. Statistical results are offered to the participants (e.g., a $t$ statistic and $p$ value), who are then asked to explicitly give or endorse various interpretations that are thought to represent participants' understanding of SST. However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication). It is unclear how well vignette studies (including ones by the present authors [@Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results. Conceptual understanding and fluency with common representations are both important, but are distinct.

@@

Using a simulated experimental task, we tested researchers' and trainees' understanding and application of SST. Our key innovation allowing us to focus on SST was to prevent---or make difficult---the use of alternative strategies. If participants have poor understanding of SST, one would predict they would perform poorly and would not be sensitive to SST-related information. In fact, scientists performed well, explicitly reported using SST strategies, and were sensitive to SST-related information. Our results strongly suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.

@@

Participants were scientists or trainees recruited via social media. Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy, where the true effect size was a randomly-assigned group difference between 0 and 1 standard deviations ($\delta$). Because the study was run around the Christmas holiday season, it was hoped that the theme would make the task more enjoyable. The numerical information for an experiment, including sample size and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}). Participants could change the sample size per group for each experiment (increasing the time required to return a result), but did not know its numerical value. Importantly, the meaning of the colors and locations was unknown to the participants, aside from the monotone relationship with the sample size and statistical evidence.

@@

Participants were also offered the opportunity to sample "random shuffle reports" that were described as the results of experiments with random assignment of elves to groups: that is, the result of experiments in which the null hypothesis was true. These results took no time to return. Participants were not told how to use these samples, only that they might use them.

@@

We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategies), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports.

@@

We determined the coding scheme and independently categorized the first 20 participant, discussing the source of disagreements. After categorizing the remaining participants, some disagreements were resolved through mutual agreement, and a discussion between the authors was had over what caused the disagreements. The remainder of the disagreements were re-coded separately, and a final round of discussion resolved the remaining disagreements. Participants' codings described in detail in Supplement B.

@@

* **Data and materials availability**: All materials, code, and data are freely available at https://github.com/richarddmorey/Morey_Hoekstra_StatCognition.

* **Author Contributions**: RDM conceptualized and designed the study in consultation with RH. RDM analysed the data and created the materials and figures. The manuscript was written by RDM and RH.

### 2020-02-05 inst/manuscript/supplementA.Rmd
small changes ahead of resubmission

```{r null-ex,echo = FALSE, out.width='100%', fig.cap="Caption"}
idx = 7
id = MoreyHoekstra2019::christmas_stats_participants$id[idx]
fn = MoreyHoekstra2019::fig_recreate_display(id, "null", letter = "N")
knitr::include_graphics(fn)
```

```{r expt-ex,echo = FALSE, out.width='100%', fig.cap="Caption"}
fn = MoreyHoekstra2019::fig_recreate_display(id, "expt", letter = "E")
knitr::include_graphics(fn)

@@

Suppose, however, that the Bayesian analyst knew about the transformation, and applied the strategy described in the previous section to reverse-engineer the transformation back to $z$ scores. The inference about $\delta$ would be dependent on their prior about the sample sizes. As emphasized by Bayesians who appeal to the Jeffreys-Lindley paradox [@Lindley:1957], the statistical support for the null hypothesis $\delta=0$ depend on the sample size such that assuming larger sample sizes guarantees evidence for the null hypothesis. The inference about $\delta$ is completely confounded with $n$: assuming larger $n$ means inferring smaller $\delta$.

@@

```{r shuffle-understanding, fig.caption="Caption"}
MoreyHoekstra2019::fig_shuffle_understanding()
```

@@

```{r confidence-null, fig.cap="Caption"}
MoreyHoekstra2019::fig_confidence(which_resp = "null")
```

```{r confidence-alt, fig.cap="Caption"}
MoreyHoekstra2019::fig_confidence(which_resp = "alt")
```

@@

```{r fig.cap="Proportion who have a 'difference' response as a function of effect size, for participants who were coded as reported 0, 1, or 2 strong significance testing strategies. Ribbons are standard errors."}

>>>>

```{r null-ex,echo = FALSE, out.width='100%', fig.cap="Selected participant's random shuffle report samples."}
idx = 7
id = MoreyHoekstra2019::christmas_stats_participants$id[idx]
fn = MoreyHoekstra2019::fig_recreate_display(id, "null", letter = "N")
knitr::include_graphics(fn)
```

```{r expt-ex,echo = FALSE, out.width='100%', fig.cap="Selected participant's experimental samples."}
fn = MoreyHoekstra2019::fig_recreate_display(id, "expt", letter = "E")
knitr::include_graphics(fn)

@@

Suppose, however, that the Bayesian analyst knew about the transformation, and applied the strategy described in the previous section to reverse-engineer the transformation back to $z$ scores. The inference about $\delta$ would be dependent on their prior about the sample sizes. As emphasized by Bayesians who appeal to the Jeffreys-Lindley paradox [@Lindley:1957], the statistical support for the null hypothesis $\delta=0$ depends on the sample size such that assuming larger sample sizes guarantees evidence for the null hypothesis. The inference about $\delta$ is completely confounded with $n$: assuming larger $n$ means inferring smaller $\delta$.

@@

```{r shuffle-understanding, fig.caption="Frequencies of responses to the initial question, \"Do you understand why the random shuffle reports could be useful?\""}
MoreyHoekstra2019::fig_shuffle_understanding()
```

@@

```{r confidence-null, fig.cap="Reported confidence ratings (vertical axis) in their response for participants who judged that they either could not detect a difference between the groups, or that the groups were the same, by true effect size (horizontal axis). Columns with very vew participants in them (when the effect size was large, but they did not respond that there was a difference) are faded to indicate lack of trustworthiness of the frequencies."}
MoreyHoekstra2019::fig_confidence(which_resp = "null")
```

```{r confidence-alt, fig.cap="Reported confidence ratings (vertical axis) in their response for participants who judged that they either could not detect a difference between the groups, or that the groups were the same, by true effect size (horizontal axis)."}
MoreyHoekstra2019::fig_confidence(which_resp = "alt")
```

@@

```{r fig.cap="Proportion who have a \"difference\" response as a function of effect size, for participants who were coded as reported 0, 1, or 2 strong significance testing strategies. Ribbons are standard errors."}
